---
title: "FE 582 Assignment Problem2"
author: "Mugdha"
date: "10/17/2020"
output: pdf_document
---

## Problem 2 
**The data provided in the files contains several quantitative and categorical variables associatewith each ticker. Please select a subset of 100 tickers from each file and use data for a specific year (ex: 2013). Use a small number of quantitative variables (10 or 12) out of ~76 columns available (example: After Tax ROE, Cash Ratio, Current Ratio, Operating Margin, Pre-Tax Margin, Pre-Tax ROE, Profit Margin, Quick Ratio, Total Assets, Total Liabilities, Earnings Per Share, etc…).The categorical variables available are GICS Sector, GICS Sub Industry, and possibly HQ Address (although this is sparse data for the 100 tickers subset selected).Next, you have to apply several distance and similarity functions to find the extreme values for distance and similarities between the subset of tickers that you chose. For each of the following cases, please define the function that allows you to calculate the quantity required, calculate the values for all ticker pairs, and rank the pairs by calculated value of distance or similarity, and report the top and bottom 10 values for each case:**  

**\textcolor{red}{Solution:}**  

```{r}
tickers<- read.csv("securities.csv")
fundamentals<- read.csv("fundamentals.csv")
#Taking only 2 categories from securities
tickers$GICS.Sector<- as.numeric(as.factor(tickers$GICS.Sector))
tickers$GICS.Sub.Industry<- as.numeric(as.factor(tickers$GICS.Sub.Industry))

#Data frame of ticker and 2 categories : GICS Sector and GICS Sub Industry
tickersfinal<-data.frame(tickers$Ticker.symbol,tickers$GICS.Sector,tickers$GICS.Sub.Industry)
names(tickersfinal)<-c("Ticker.symbol","GICS.Sector",
                       "GICS.Sub.Industry")

#Selecting year 2012
Dataset<- subset(fundamentals,format(as.Date(fundamentals$Period.Ending),"%Y")==2012)

#Creating a data frame of 9 quantitative features from fundamentals:Ticker symbol,Cash Ratio,Current Ratio,After.Tax.ROE,Gross.Margin,Profit.Margin,Pre.Tax.ROE,Total.Assets,Total.Liabilities
FDataset<- data.frame(Dataset$Ticker.Symbol,Dataset$Cash.Ratio,Dataset$Current.Ratio,Dataset$After.Tax.ROE,Dataset$Gross.Margin,Dataset$Profit.Margin,Dataset$Pre.Tax.ROE,Dataset$Total.Assets,Dataset$Total.Liabilities)
                      
#Taking subset of 100 from data of 2012 year
Finaldataset<- head(FDataset,100)

names(Finaldataset)<- c("Ticker.symbol","Cash.Ratio","Current.Ratio","After.Tax.ROE","Gross.Margin","Profit.Margin","Pre.Tax.ROE","Total.Assets","Total.Liabilities")

#Merging data set of both quantitative and categorical data (11 features)to form a Final Data which I will be working on for all the problems: 
Finaldata<-merge(Finaldataset,tickersfinal,by="Ticker.symbol",all.x=TRUE)

#Removing NA values from dataset
Finaldata<- na.omit(Finaldata)

```
**a,b,c,d)Calculating Lp norms for p=1,2,3,10**

**\textcolor{red}{Solution:}**  
```{r}
lpnorm <- function(feature_data,rownum,p)
{
  #Selecting only 1 to 7 features for now as they are quantitative and 10 and 11 are categorical.Not taking 8 and 9 quantitative features as they are large 
  feature_data<- feature_data[1:7]
  ldist =c()
  for(j in 1:nrow(feature_data))
  {
    result<-(sum(abs(feature_data[rownum,2:7]-
            feature_data[j,2:7])^p))^(1/p)
    templ1dist = c(feature_data[rownum,1],feature_data[j,1],result)
    ldist<-append(ldist, templ1dist)
  }
  #Output in the form of Ticker1, Ticker 2 and Distance
  matrix1 <- matrix(ldist, ncol=3, byrow=TRUE)
  df <- as.data.frame(matrix1, stringsAsFactors=FALSE)
  df <- na.omit(df)
  df$V3 <- as.numeric(as.character(df$V3))
  dfSorted <-df[order(-df$V3),]
  names(dfSorted)<- c("Ticker1","Ticker2","Distance")
  #Top 10 
  print(head(dfSorted, 10))
  #Bottom 10
  print(tail(dfSorted, 10))
}
#Distance of ticker 1 (Row=1) with other tickers, p=1
lpnorm(Finaldata,1,1)
#Row=1 ,p=2
lpnorm(Finaldata,1,2)
#Row=1 ,p=3
lpnorm(Finaldata,1,3)
#Row=1 ,p=10
lpnorm(Finaldata,1,10)
```
**e)Minkovski distance (assign different weights for the feature components in the Lp-norm based on your assessment on the importance of the features)**

**\textcolor{red}{Solution:}**  
```{r}
minkovskiDist <- function(feature_data,rownum,p)
{
  #Selecting only 1 to 9 features as they are quantitative    and 10 and 11 are categorical
  feature_data<- feature_data[1:9]
  l1dist =c()
  for(j in 1:nrow(feature_data))
  {
    
    result<-(sum(abs(feature_data[rownum,2:7]-feature_data[j,2:7])^p))
    
    #Assigning weight to Assets and Liabilities features
    resultTotalAssetLiab<-(sum((1/1000000000) * (abs(feature_data[rownum,8:9]-feature_data[j,8:9])^p)))
    finalResult<-(sum(result, resultTotalAssetLiab))^(1/p)
    
    templ1dist = c(feature_data[rownum,1], feature_data[j,1], finalResult)
    l1dist<-append(l1dist, templ1dist)
  }
  #Output in the form of Ticker1, Ticker 2 and Distance
  m1 <- matrix(l1dist, ncol=3, byrow=TRUE)
  df <- as.data.frame(m1, stringsAsFactors=FALSE)
  #print(df)
  df <- na.omit(df)
  df$V3 <- as.numeric(as.character(df$V3))
  dfSorted <-df[order(-df$V3),]
  #print(dfSorted)
  names(dfSorted)<- c("Ticker1","Ticker2","Distance")
  print(head(dfSorted, 10))
  print(tail(dfSorted, 10))
}
#Minkovski distance for row=1(First row) with other tickers
minkovskiDist(Finaldata,1,1)
```
**Assets and Liabilities in the data set had values in billions which resulted in large distances. Hence assigned weights to these features**

**f)Match-Based Similarity Computation**

**\textcolor{red}{Solution:}**  
```{r}
matchSim <- function(feature_data,rownum,p)
{
  #Selecting 1 to 9 as those are the 9 quantitative features of the data
  feature_data<- feature_data[1:9]
  matchdist =c()
  for(j in 1:nrow(feature_data))
  {
    tempMatchDist = c()
    for(k in 2:ncol(feature_data)) {
      max = max(feature_data[,k])
      min = min(feature_data[,k])
     
  # calculate the bucket ranges using bucket size as 3
     bucketRange = round(((max-min)/3),0)
  # create buckets
     feature_data$buck= cut(feature_data[,k],c(min,bucketRange,bucketRange*2,max),
                  labels=c("Bucket1","Bucket2","Bucket3"),
                  include.lowest = TRUE)
  # check if the feature belongs to same bucket
      if(feature_data[rownum,10] == feature_data[j,10]){
        # find min and max of the bucket
        minBuck=c()
        maxBuck=c()
        if(feature_data[rownum,10] == "Bucket1") {
          minBuck = min  
          maxBuck = bucketRange
        } else if(feature_data[rownum,10] == "Bucket2") {
          minBuck = bucketRange+1  
          maxBuck = bucketRange*2
        } else {
          minBuck = bucketRange*2+1 
          maxBuck = max
        }
        # compute the expression
        result = (1-abs(feature_data[rownum,k]-feature_data[j,k])/(maxBuck-minBuck))^p
        
        # add to tempMatchDist
        tempMatchDist = append(tempMatchDist,result)
      }
      # removing temporary bucket column
      feature_data$buck <- NULL
    }
    # add tickers and tempMatchDist to matchDist
    finalResult = c(feature_data[rownum,1], feature_data[j,1], sum(tempMatchDist)^(1/p))
    matchdist = append(matchdist, finalResult)
  }
  # sorting and printing
   #Output in the form of Ticker1, Ticker 2 and Distance
  m1 <- matrix(matchdist, ncol=3, byrow=TRUE)
  df <- as.data.frame(m1, stringsAsFactors=FALSE)
  df <- na.omit(df)
  df$V3 <- as.numeric(as.character(df$V3))
  dfSorted <-df[order(-df$V3),]
  names(dfSorted)<- c("Ticker1","Ticker2","Match Based Similarity")
  #Top 10
  print(head(dfSorted, 10))
  #Bottom 10
  print(tail(dfSorted, 10))
}
#Match based similarity for row=1 (First row) and p=2 with other tickers
matchSim(Finaldata,1,2)
```

**g)Mahalanobis distance**

**\textcolor{red}{Solution:}**  
```{r}
mahalanoDist <- function(feature_data,rownum){
  mahaDist <- c()
  for(j in 1:nrow(feature_data)){
    feature_data<- na.omit(feature_data)
    #Calculating Covariance of Feature Data
    #Selecting only first 7 quantitative features 
    CovMat <- cov(feature_data[2:7])
    matrix_S <-as.matrix(feature_data[rownum,2:7]-
                         feature_data[j,2:7])[1,]     
    vector1<-(as.vector(matrix_S))
    #Calculating the expression
    dmahal <- vector1 %*% solve(as.matrix(CovMat)) 
    finalmaha <- dmahal * t(matrix_S) 
    mahafinal<-(sum(finalmaha))
   tempmahafinal<-c(feature_data[rownum,1], feature_data[j,1], mahafinal)
    mahaDist <- append(mahaDist,tempmahafinal)
  }
  #Output in the form of Ticker1, Ticker 2 and Distance
  m1 <- matrix(mahaDist, ncol=3, byrow=TRUE)
  df <- as.data.frame(m1, stringsAsFactors=FALSE)
  df <- na.omit(df)
  df$V3 <- as.numeric(as.character(df$V3))
  dfSorted <-df[order(-df$V3),]
  names(dfSorted)<- c("Ticker1","Ticker2","Distance")
  #Top 10
  print(head(dfSorted, 10))
  #Bottom 10
  print(tail(dfSorted, 10))

}
#Mahalanolobis Distance for row=1 (First row) with other tickers
mahalanoDist(Finaldata,1)

```
**h)Similarity: overlap measure**

**\textcolor{red}{Solution:}**  
```{r}
olap<- function(feature_data,rownum)
{
  #Taking categorical features Sector and Sub sector from the data set
  feature_data<- feature_data[c(1,10:11)]
  feature_data<- na.omit(feature_data)
  opdist <- c()
  for(j in 1:nrow(feature_data))
  {
    result = c()
    #Checking if same category for both the features
    if(feature_data[rownum,2]==feature_data[j,2])
    {
       result = append(result,1)
    }
    if(feature_data[rownum,3]==feature_data[j,3])
    {
       result = append(result,1)  
    }
    tempOpDist =  c(feature_data[rownum,1],feature_data[j,1],sum(result))  
    opdist<- append(opdist, tempOpDist)
  }
   #Output in the form of Ticker1, Ticker 2 and Overlap
  m1 <- matrix(opdist, ncol=3, byrow=TRUE)
  df <- as.data.frame(m1, stringsAsFactors=FALSE)
  df <- na.omit(df)
  df$V3 <- as.numeric(as.character(df$V3))
  dfSorted <-df[order(-df$V3),]
  names(dfSorted)<- c("Ticker1","Ticker2","Overlap Measure")
  #Top 10
  print(head(dfSorted, 10))
  #Bottom 10
  print(tail(dfSorted, 10))
}

#Overlap measure for row=7 with other tickers
olap(Finaldata,7)
```
**i)Similarity: inverse frequency**

**\textcolor{red}{Solution:}**  
```{r}
invf<- function(feature_data,rownum)
{   #Taking categorical features Sector and Sub sector from     the data set
  feature_data<- feature_data[c(1,10:11)]
  iofDist <- c()
  for(j in 1:nrow(feature_data))
  {
    result = c()
    #Checking if same category
    if(feature_data[rownum,2]==feature_data[j,2])
    {
    #Calculating the fraction of records for same category feature
      fractionDF = feature_data[feature_data$GICS.Sector == feature_data[rownum,2],]
      p = length(fractionDF$GICS.Sector) / 100
      # Calculating the expression
      result = append(result,(1/p)^2)
    }
    if(feature_data[rownum,3]==feature_data[j,3])
    {
    #Calculating the fraction of records for same categoryfeature
      fractionDF = feature_data[feature_data$GICS.Sector == feature_data[rownum,3],]
      p = length(fractionDF$GICS.Sector) / 100
       # Calculating the expression
      result = append(result,(1/p)^2)
    }
    tempIofDist = c(feature_data[rownum,1],feature_data[j,1],sum(result))  
    iofDist<- append(iofDist, tempIofDist)
  }
   #Output in the form of Ticker1, Ticker 2 and Similarity
  m1 <- matrix(iofDist, ncol=3, byrow=TRUE)
  df <- as.data.frame(m1, stringsAsFactors=FALSE)
  df <- na.omit(df)
  df$V3 <- as.numeric(as.character(df$V3))
  dfSorted <-df[order(-df$V3),]
  names(dfSorted)<- c("Ticker1","Ticker2","Similarity")
  #Top 10
  print(head(dfSorted, 10))
  #Bottom 10
  print(tail(dfSorted, 10))
}
#Inverse occurrence frequency for row=1 (First row) with other tickers
invf(Finaldata,1)
```
**j)Similarity: Goodall**

**\textcolor{red}{Solution:}**  
```{r}
gooDallSim<- function(feature_data,rownum)
{ 
  feature_data<- feature_data[c(1,10:11)]
  iofDist <- c()
  for(j in 1:nrow(feature_data))
  {
    result = c()
    if(feature_data[rownum,2]==feature_data[j,2])
    {
      fractionDF = feature_data[feature_data$GICS.Sector == feature_data[rownum,2],]
    #Calculating the fraction of records for same category feature
      p = length(fractionDF$GICS.Sector) / 100
       # Calculating the expression
      result = append(result,(1-p^2))
    }
    if(feature_data[rownum,3]==feature_data[j,3])
    {
      fractionDF = feature_data[feature_data$GICS.Sector == feature_data[rownum,3],]
    #Calculating the fraction of records for same category feature
      p = length(fractionDF$GICS.Sector) / 100
        #Calculating the expression
      result = append(result,(1-p^2))
    }
    tempIofDist = c(feature_data[rownum,1],feature_data[j,1],sum(result))  
    iofDist<- append(iofDist, tempIofDist)

  }
  #Output in the form of Ticker1, Ticker 2 and Similarity
  m1 <- matrix(iofDist, ncol=3, byrow=TRUE)
  df <- as.data.frame(m1, stringsAsFactors=FALSE)
  df <- na.omit(df)
  df$V3 <- as.numeric(as.character(df$V3))
  dfSorted <-df[order(-df$V3),]
  names(dfSorted)<- c("Ticker1","Ticker2","Similarity")
  #Top 10
  print(head(dfSorted, 10))
  #Bottom 10
  print(tail(dfSorted, 10))
}

#Goodall Measure for row=1 (First row) with other tickers
gooDallSim(Finaldata,2)

```
**k)Overall similarity between tickers by using mixed type data (choose a lambda value for calculation)**

**\textcolor{red}{Solution:}**  

```{r}
overallsim <- function(feature_data,rownum,p,lmda)
{
  #calculating match based similarity for NumSim
  overallSimilarities = c()
  for(j in 1:nrow(feature_data))
  {
    tempMatchDist = c()
    #Taking columns 2 to 9 as they are quantitative features
    for(k in 2:(ncol(feature_data)-2)) {
      max = max(feature_data[,k])
      min = min(feature_data[,k])
      # calculate the bucket ranges
      bucketRange = round(((max-min)/3),0)
      # create buckets
      feature_data$buck= cut(feature_data[,k], c(min,bucketRange,bucketRange*2,max),
                             labels = c("Bucket1","Bucket2","Bucket3"),
                             include.lowest = TRUE)

      # check if features belongs to same bucket
      if(feature_data[rownum,12] == feature_data[j,12]){
        # find min and max of the bucket
        minBuck=c()
        maxBuck=c()
        if(feature_data[rownum,12] == "Bucket1") {
          minBuck = min
          maxBuck = bucketRange
        } else if(feature_data[rownum,12] == "Bucket2") {
          minBuck = bucketRange+1
          maxBuck = bucketRange*2
        } else {
          minBuck = bucketRange*2+1
          maxBuck = max
        }
        # compute the expression
        result = (1-abs(feature_data[rownum,k]-feature_data[j,k])/(maxBuck-minBuck))^p

        # add to tempMatchDist
        tempMatchDist = append(tempMatchDist,result)
      }
      # removing temporary bucket column
      feature_data$buck <- NULL
    }
     # compute the expression
    numSim = sum(tempMatchDist)^(1/p) * lmda

    # Calculate CatSim using overlap measure
    opdist <- c()
    #Taking features 10 and 11 as they are categorical features
    if(feature_data[rownum,10]==feature_data[j,10])
    {
      opdist = append(opdist,1)
    }
    if(feature_data[rownum,11]==feature_data[j,11])
    {
      opdist = append(opdist,1) 
    }
     # compute the expression
    catSim <- sum(opdist) * (1-lmda)
    
    overallSimilarity = c(feature_data[rownum,1], feature_data[j,1], sum(numSim,catSim))
    overallSimilarities = append(overallSimilarities, overallSimilarity)
  }
  # sorting and printing
   #Output in the form of Ticker1, Ticker 2 and Similarity
  m1 <- matrix(overallSimilarities, ncol=3, byrow=TRUE)
  df <- as.data.frame(m1, stringsAsFactors=FALSE)
  df <- na.omit(df)
  df$V3 <- as.numeric(as.character(df$V3))
  dfSorted <-df[order(-df$V3),]
  names(dfSorted)<- c("Ticker1","Ticker2","Similarity")
  #Top 10
  print(head(dfSorted, 10))
  #Bottom 10
  print(tail(dfSorted, 10))
}
#Calculating overall similarity by taking row=1, p=2 and lambda value as 0.25
overallsim(Finaldata,1,2,0.25)

```
**l)Overall normalized similarity between tickers by using mixed type data (choose a lambda value for calculation)**


**\textcolor{red}{Solution:}**  
```{r}
overallSimNormalized <- function(feature_data,rownum,p,lmda)
{
  #calculating match based similarity for NumSim
  overallSimilarities = c()
  for(j in 1:nrow(feature_data))
  {
    tempMatchDist = c()
    for(k in 2:(ncol(feature_data)-2)) {
      max = max(feature_data[,k])
      min = min(feature_data[,k])
      # calculate the bucket ranges
      bucketRange = round(((max-min)/3),0)
      # create buckets
      feature_data$buck= cut(feature_data[,k], c(min,bucketRange,bucketRange*2,max),
                             labels = c("Bucket1","Bucket2","Bucket3"),
                             include.lowest = TRUE)
      
      # check if features belongs to same bucket
      if(feature_data[rownum,12] == feature_data[j,12]){
        # find min and max of the bucket
        minBuck=c()
        maxBuck=c()
        if(feature_data[rownum,12] == "Bucket1") {
          minBuck = min
          maxBuck = bucketRange
        } else if(feature_data[rownum,12] == "Bucket2") {
          minBuck = bucketRange+1
          maxBuck = bucketRange*2
        } else {
          minBuck = bucketRange*2+1
          maxBuck = max
        }
        # compute the expression
        result = (1-abs(feature_data[rownum,k]-feature_data[j,k])/(maxBuck-minBuck))^p
        
        # add to tempMatchDist
        tempMatchDist = append(tempMatchDist,result)
      }
      # removing temporary bucket column
      feature_data$buck <- NULL
    }
    # compute the expression
    overallNumSim = sum(tempMatchDist)^(1/p)
    # Calculate CatSim using overlap measure
    opdist <- c()
    if(feature_data[rownum,10]==feature_data[j,10])
    {
      opdist = append(opdist,1)
    }
    if(feature_data[rownum,11]==feature_data[j,11])
    {
      opdist = append(opdist,1) 
    }
    overallCatSim = sum(opdist)
    overallSimilarity = c(feature_data[rownum,1], feature_data[j,1], overallNumSim, overallCatSim)
    overallSimilarities = append(overallSimilarities, overallSimilarity)
  }
  # sorting and printing
   #Output in the form of Ticker1, Ticker 2 and Similarity
  m1 <- matrix(overallSimilarities, ncol=4, byrow=TRUE)
  df <- as.data.frame(m1, stringsAsFactors=FALSE)
  #print(df)
  df <- na.omit(df)
  df$V3 <- as.numeric(as.character(df$V3))
  df$V4 <- as.numeric(as.character(df$V4))
  
  sigmaNum = sd(df$V3)
  sigmaCat = sd(df$V4)
  #calculating the sigmaNum and sigmaCat values for quantitative and categorical features
  df$V5 <- (df$V3 * (1/sigmaNum) * lmda) + (df$V4 * (1/sigmaCat) * (1-lmda))
  df$V3<-NULL
  df$V4<-NULL
  df$V5 <- as.numeric(as.character(df$V5))
  dfSorted <-df[order(-df$V5),]
  # print(dfSorted)
  names(dfSorted)<- c("Ticker1","Ticker2","Similarity")
  print(head(dfSorted, 10))
  print(tail(dfSorted, 10))
}
#Calculating overall normalized similarity by taking row=1, p=2 and lambda value as 0.25
overallSimNormalized(Finaldata,1,2,0.25)

```